Design and Implementation of a Multithreaded Virtual Machine for Executing Linear Logic Programs

### 2014-07-25 paper.tex
Update.

\title{Design and Implementation of a Multithreaded Virtual Machine for Executing Linear Logic Programs}

>>>>

\toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from Permissions@acm.org.

PPDP '14, September 08 - 10 2014, Canterbury, United Kingdom
Copyright 2014 ACM 978-1-4503-2947-7/14/09\$15.00.
http://dx.doi.org/10.1145/2643135.2643150 }
\title{Design and Implementation of a Multithreaded Virtual Machine for Executing Linear Logic Programs}

### 2014-07-25 virtual_machine.tex
Update.

A key goal of our design is to keep the threads as busy as possible
and to reduce inter-thread communication. When the VM starts, it reads
the byte-code file and starts all threads. Initially, the VM will
partition the application graph of $N$ nodes into $T$ subgraphs (the
number of threads) and then each thread will work on their own
subgraph. Reduction of inter-thread communication is achieved by first
ordering the node addresses present in the code in such a way that
closer nodes are clustered together and then partitioning them to
threads. During compilation, we take note of predicates that are used
in rules with communication (arguments with type \emph{node}) and
then build a graph of nodes from the program's axioms. The nodes of
the graph are then ordered by using a breadth-first search algorithm
that changes the nodes of addresses to the domain $[0, n[$, where $n$
is the number of nodes. Once the VM starts, we simply partition the
range $[0, n[$ into $T$ subgraphs.

@@

Figure~\ref{fig:overview} presents the layout of our virtual machine
for a program with six nodes and two running threads. Each thread
space includes the nodes owned by the thread (the dotted arrows
represent the edges between nodes) and a \emph{Work Queue}, a single
linked list containing \emph{active nodes}, i.e., nodes that have new
facts to process. Initially, the \emph{Work Queue} is filled with all
the nodes of the thread in order to derive the
axioms. Figure~\ref{fig:overview} also illustrates the internal
structure layout of a node, which includes: the database of linear
facts (\emph{Linear DB}); the database of persistent facts
(\emph{Persistent DB}); the rule matching structures (\emph{Rule
Engine}); and a auxiliary buffer for storing intermediate facts
coming from other threads (\emph{Fact Buffer}).

@@

{\footnotesize
\begin{Verbatim}
// rules
a, e(1) -o b.
a -o c.
b -o d.
e(0) -o f.
c -o e(1).

@@

Note that every node in the program has the same set of data
structures presented in Fig.~\ref{fig:rule_engine}. We use 32 bits
integers to implement bitmaps and an array of 16 bits integers to
count facts, resulting in $32 + 2P$ bytes per node, where $P$ is the
number of predicates.

@@

>>>>

A key goal of our design is to keep the threads as busy as possible
and to reduce inter-thread communication. When the VM starts, it reads
the byte-code file and starts all threads. Initially, the VM will
partition the application graph of $N$ nodes into $T$ subgraphs (the
number of threads) and then each thread will work on their own
subgraph. Reduction of communication between nodes in different threads
is achieved by first ordering the node addresses present in the code in such a way that
closer nodes are clustered together and then partitioning them to
threads. During compilation, we take note of predicates that are used
in rules with communication (arguments with type \emph{node}) and
then build a graph of nodes from the program's axioms. The nodes of
the graph are then ordered by using a breadth-first search algorithm
that changes the nodes of addresses to the domain $[0, n[$, where $n$
is the number of nodes. Once the VM starts, we simply partition the
range $[0, n[$ into $T$ subgraphs.

@@

Figure~\ref{fig:overview} presents the layout of our virtual machine
for a program with six nodes and two running threads. Each thread
space includes the nodes owned by the thread (the dotted arrows
represent the edges between nodes) and a \emph{Work Queue}, a single
linked list containing \emph{active nodes}, i.e., nodes that have new
facts to process. Initially, the \emph{Work Queue} is filled with all
the nodes of the thread in order to derive the
axioms. Figure~\ref{fig:overview} also illustrates the internal
structure layout of a node, which includes: the database of linear
facts (\emph{Linear DB}); the database of persistent facts
(\emph{Persistent DB}); the rule matching structures (\emph{Rule
Engine}); and an auxiliary buffer for storing intermediate facts
coming from other threads (\emph{Fact Buffer}).

@@

{\footnotesize
\begin{Verbatim}
// rules
1: a, e(1) -o b.
2: a -o c.
3: b -o d.
4: e(0) -o f.
5: c -o e(1).

@@

Note that every node in the program has the same set of data
structures presented in Fig.~\ref{fig:rule_engine}. We use 32 bits
integers to implement the 3 bitmaps and an array of 16 bits integers to
count facts.

@@

We avoid garbage collection schemes since objects are created and discarded
in very specific points of the virtual machine and our objects cannot contain circular references.
A reference counting mechanism is thus more appropriate than a parallel garbage collector which
would entail pausing the execution of the program to garbage collect all the unused objects.
